---
title: "Basic Sampling for Poisson Regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Sampling for Poisson Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: Vignette_Ref.bib
---

<style>
body {
text-align: justify}
</style>

```{r load packages and theme,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NeEDS4BigData)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(tidyr)
library(psych)
library(gam)

set.seed(1234)

# Theme for plots
Theme_special<-function(){
  theme(legend.key.width= unit(1,"cm"),
        axis.text.x= element_text(color= "black", size= 12, angle= 30, hjust=0.75),
        axis.text.y= element_text(color= "black", size= 12), 
        strip.text= element_text(colour= "black", size= 12, face= "bold"),
        panel.grid.minor.x= element_blank(),
        axis.title= element_text(color= "black", face= "bold", size= 12),
        legend.text= element_text(color= "black", size= 11),
        legend.position= "bottom",
        legend.margin= margin(0,0,0,0),
        legend.box.margin= margin(-1,-2,-1,-2)) 
}
```

# Understanding the bike sharing data

@fanaee2013bike collected data to understand the bike sharing demands under the rental and return process.
The data contains $4$ columns and has $17,379$ observations, first column is the response variable and the rest are covariates. 
We consider the covariates a) temperature ($x_1$), b) humidity ($x_2$) and c) wind speed ($x_3$) to model the response, the number of bikes rented hourly.
The covariates are scaled to be mean of $\mu=0$ and variance of $\sigma^2=1$.
For the given data subsampling methods are implemented assuming the main effects model can describe the data.

```{r Exploring data}
Original_Data<-cbind(Bike_sharing[,1],1,Bike_sharing[,-1])
colnames(Original_Data)<-c("Y",paste0("X",0:ncol(Original_Data[,-c(1,2)])))

# Scaling the covariate data
for (j in 3:5) {
  Original_Data[,j]<-scale(Original_Data[,j])
}
head(Original_Data) %>% 
  kable(format = "html",
        caption = "First five observations of the bike sharing data.")

# Setting the subsample sizes
N<-nrow(Original_Data); M<-250; k<-seq(6,18,by=3)*100; rep_k<-rep(k,each=M)
```

Based on this model the methods random sampling, leverage sampling [@ma2014statistical;@ma2015leveraging], $A$- and $L$-optimality subsampling [@ai2021optimal;@yao2021review] and $A$-optimality subsampling with response constraint [@zhang2021optimal] are implemented on the big data.

```{r define subsampling methods}
# define colors, shapes, line types for the methods
Method_Names<-c("A-Optimality","L-Optimality","A-Optimality MC",
                "Shrinkage Leverage","Basic Leverage","Unweighted Leverage",
                "Random Sampling")
Method_Color<-c("green","green4","palegreen",
                "red","darkred","maroon","black")
Method_Shape_Types<-c(rep(8,3),rep(4,3),16)
Method_Line_Types<-c(rep("twodash",3),rep("dotted",3),"solid")
```

The obtained subsamples and their respective model parameter estimates are over $M=100$ simulations across different subsample sizes $k=(600,\ldots,1500)$. 
We set the initial subsample size of $r1=300$ for the methods that requires a random sample. 
From the final subsamples, model parameters of the assume model are estimated.
These estimated model parameters are compared with the estimated model parameters of the full big data through the mean squared error $MSE(\tilde{\beta}_k,\hat{\beta})=\frac{1}{MJ}\sum_{i=1}^M \sum_{j=1}^J(\tilde{\beta}_{k,j} - \hat{\beta}_j)^2$. 
Here, $\tilde{\beta}_k$ is the estimated model parameters from the subsample of size $k$ and $\hat{\beta}$ is the estimated model parameters from the full big data, while $j$ is index of the model parameter.  
Below is the code for implementing these subsampling methods.

## Random sampling

```{r Random sampling}
# Random sampling
Final_Beta_RS<-matrix(nrow = length(k)*M,ncol = ncol(Original_Data[,-1])+1)
for (i in 1:length(rep_k)) {
  Temp_Data<-Original_Data[sample(1:N,rep_k[i]),]
  glm(Y~.-1,data=Temp_Data,family="poisson")->Results
  Final_Beta_RS[i,]<-c(rep_k[i],coefficients(Results))
  if(i==length(rep_k)){print("All simulations completed for random sampling")}
}
Final_Beta_RS<-cbind.data.frame("Random Sampling",Final_Beta_RS)
colnames(Final_Beta_RS)<-c("Method","Subsample",paste0("Beta",0:3))
```

## Leverage sampling

```{r Leverage sampling}
# Leverage sampling 
## we set the shrinkage value of alpha=0.9
NeEDS4BigData::LeverageSampling(r=rep_k,Y=as.matrix(Original_Data[,1]),
                                X=as.matrix(Original_Data[,-1]),N=N,
                                alpha = 0.9,family = "poisson")->Results
Final_Beta_LS<-Results$Beta_Estimates
colnames(Final_Beta_LS)<-c("Method","Subsample",paste0("Beta",0:3))
```

## A- and L-optimality subsampling

```{r A- and L-optimality subsampling}
# A- and L-optimality subsampling for GLM
NeEDS4BigData::ALoptimalGLMSub(r1=300,r2=rep_k,
                               Y=as.matrix(Original_Data[,1]),
                               X=as.matrix(Original_Data[,-1]),
                               N=N,family = "poisson")->Results
Final_Beta_ALoptGLM<-Results$Beta_Estimates
colnames(Final_Beta_ALoptGLM)<-c("Method","Subsample",paste0("Beta",0:3))
```

## A-optimality subsampling with measurement constraints

```{r A-optimality subsampling with measurement constraints}
# A-optimality subsampling without response
NeEDS4BigData::AoptimalMCGLMSub(r1=300,r2=rep_k,
                                Y=as.matrix(Original_Data[,1]),
                                X=as.matrix(Original_Data[,-1]),
                                N=N,family="poisson")->Results
Final_Beta_AoptMCGLM<-Results$Beta_Estimates
Final_Beta_AoptMCGLM$Method<-rep("A-Optimality MC",nrow(Final_Beta_AoptMCGLM))
colnames(Final_Beta_AoptMCGLM)<-c("Method","Subsample",paste0("Beta",0:3))
```

## Summary

```{r summarising results}
# Summarising Results of Scenario one
Final_Beta<-rbind(Final_Beta_RS,Final_Beta_LS,
                  Final_Beta_ALoptGLM,Final_Beta_AoptMCGLM)

# Obtaining the model parameter estimates for the full big data model
glm(Y~.-1,data = Original_Data,family="poisson")->All_Results
coefficients(All_Results)->All_Beta
matrix(rep(All_Beta,by=nrow(Final_Beta)),nrow = nrow(Final_Beta),
       ncol = ncol(Final_Beta[,-c(1,2)]),byrow = TRUE)->All_Beta

remove(Final_Beta_RS,Final_Beta_LS,Final_Beta_ALoptGLM,
       Final_Beta_AoptMCGLM,Temp_Data,Results,All_Results)
```

The mean squared error is plotted below.

```{r First scenario model parameter estimates,fig.width=7,fig.height=12,fig.align='center',fig.cap="Mean squared error for a) all the subsampling methods and b) without unweighted leverage sampling."}
# Obtain the mean squared error for the model parameter estimates
MSE_Beta<-data.frame("Method"=Final_Beta$Method,
                     "Subsample"=Final_Beta$Subsample,
                     "MSE"=rowSums((All_Beta - Final_Beta[,-c(1,2)])^2)) 

MSE_Beta$Method<-factor(MSE_Beta$Method,levels = Method_Names,labels = Method_Names)

# Plot for the mean squared error with all methods
ggplot(MSE_Beta |> dplyr::group_by(Method,Subsample) |> 
       dplyr::summarise(MSE=mean(MSE), .groups = 'drop'),
       aes(x=factor(Subsample),y=MSE,color=Method,group=Method,
           linetype=Method,shape=Method))+
  geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
  scale_color_manual(values = Method_Color)+
  scale_linetype_manual(values=Method_Line_Types)+
  scale_shape_manual(values = Method_Shape_Types)+
  theme_bw()+guides(colour = guide_legend(nrow = 3))+Theme_special()->p1

# Plot for the mean squared error with all methods except unweighted leverage
ggplot(MSE_Beta[MSE_Beta$Method != "Unweighted Leverage",] |> 
        dplyr::group_by(Method,Subsample) |> 
        dplyr::summarise(MSE=mean(MSE), .groups = 'drop'),
       aes(x=factor(Subsample),y=MSE,color=Method,group=Method,
           linetype=Method,shape=Method))+
  geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
  scale_color_manual(values = Method_Color[-7])+
  scale_linetype_manual(values=Method_Line_Types[-7])+
  scale_shape_manual(values = Method_Shape_Types[-7])+
  theme_bw()+guides(colour = guide_legend(nrow = 3))+Theme_special()->p2

#save(Final_Beta,All_Beta,MSE_Beta,p1,p2,file="Poisson_Regression_Scenario_1.RData")
#load("Poisson_Regression_Scenario_1.RData")
ggarrange(p1,p2,nrow = 2,ncol = 1,labels = "auto")
```

## References
