---
title: "Basic Sampling for Logistic Regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Sampling for Logistic Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r load packages and theme,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NeEDS4BigData)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(tidyr)
library(psych)
library(gam)

# Theme for plots
Theme_special<-function(){
  theme(legend.key.width= unit(1,"cm"),
        axis.text.x= element_text(color= "black", size= 12, angle= 30, hjust=0.75),
        axis.text.y= element_text(color= "black", size= 12), 
        strip.text= element_text(colour= "black", size= 12, face= "bold"),
        panel.grid.minor.x= element_blank(),
        axis.title= element_text(color= "black", face= "bold", size= 12),
        legend.text= element_text(color= "black", size= 11),
        legend.position= "bottom",
        legend.margin= margin(0,0,0,0),
        legend.box.margin= margin(-1,-2,-1,-2)) 
}
```

# Understanding the skin segmentation data

The data contains $4$ columns and has $245,057$ observations, first column is the response variable and the rest are covariates, however we only use $25\%$ of the data for this explanation.
Aim of the logistic regression model is to classify if images are skin or not based on the a) Red, b) Green and c) Blue colour data.
Skin presence is denoted as one and skin absence is denoted as zero.
Each colour vector is scaled to have a mean of zero and a variance of one (initial range was between 0−255).
For the given data subsampling methods are implemented assuming the main effects model can describe the data.

```{r Exploring data}
indexes<-sample(1:nrow(Skin_segmentation),nrow(Skin_segmentation)*0.25)
Original_Data<-cbind(Skin_segmentation[indexes,1],1,Skin_segmentation[indexes,-1])
colnames(Original_Data)<-c("Y",paste0("X",0:ncol(Original_Data[,-c(1,2)])))

# Scaling the covariate data
for (j in 3:5) {
  Original_Data[,j]<-scale(Original_Data[,j])
}
head(Original_Data) %>% 
  kable(format = "html",
        caption = "First five observations of the skin segmentation data.")

# Setting the subsample sizes
N<-nrow(Original_Data); M<-200; k<-c(6:15)*100; rep_k<-rep(k,each=M)
```

We focus on assuming the main effects model with the intercept can appropriately describe the big data.
Based on this model the methods random sampling, leverage sampling, local case control sampling, $A$- and $L$-optimality subsampling and $A$-optimality subsampling with response constraint are implemented on the big data.

```{r Define the methods}
# define colors, shapes, line types for the methods
Method_Names<-c("A-Optimality","L-Optimality","A-Optimality MC",
                "Local case control sampling","Shrinkage Leverage",
                "Basic Leverage","Unweighted Leverage","Random sampling")
Method_Color<-c("green","green4","palegreen",
                "maroon","red","darkred","firebrick","black")
Method_Shape_Types<-c(rep(8,3),rep(4,4),16)
Method_Line_Types<-c(rep("twodash",3),rep("dotted",4),"solid")
```

The obtained subsamples and their respective model parameter estimates are over $M=100$ simulations across different subsample sizes $k=(600,\ldots,1500)$.
We set the initial subsample size of $r1=300$ for the methods that requires a random sample. 
From the final subsamples their respective model parameters are estimated. 
These estimated model parameters are compared with the estimated model parameters of the full big data through the mean squared error $MSE(\tilde{\beta}_k,\hat{\beta})=\sum_{i=1}^M (\tilde{\beta}_k - \hat{\beta})^2/M$. 
Here, $\tilde{\beta}_k$ is the estimated model parameters from the subsample of size $k$ and $\hat{\beta}$ is the estimated model parameters from the full big data. 
Below is the code for this implementation.

## Random sampling

```{r Random sampling}
# Random sampling
Final_Beta_RS<-matrix(nrow = length(k)*M,ncol = ncol(Original_Data[,-1])+1)
for (i in 1:length(rep_k)) {
  Temp_Data<-Original_Data[sample(1:N,rep_k[i]),]
  glm(Y~.-1,data=Temp_Data,family="binomial")->Results
  Final_Beta_RS[i,]<-c(rep_k[i],coefficients(Results))
  if(i==length(rep_k)){print("All simulations completed for random sampling")}
}
Final_Beta_RS<-cbind.data.frame("Random sampling",Final_Beta_RS)
colnames(Final_Beta_RS)<-c("Method","Subsample",paste0("Beta",0:3))
```

## Leverage sampling

```{r Leverage sampling}
# Leverage sampling
## we set the shrinkage value of alpha=0.9
NeEDS4BigData::LeverageSampling(r=rep_k,Y=as.matrix(Original_Data[,1]),
                                X=as.matrix(Original_Data[,-1]),N=N,
                                alpha = 0.9,family = "logistic")->Results
Final_Beta_LS<-Results$Beta_Estimates
colnames(Final_Beta_LS)<-c("Method","Subsample",paste0("Beta",0:3))
```

## Local case control sampling

```{r local case control sampling}
# Local case control sampling
NeEDS4BigData::LCCsampling(r1=300,r2=rep_k,
                           Y=as.matrix(Original_Data[,1]),
                           X=as.matrix(Original_Data[,-1]),
                           N=N)->Results
Final_Beta_LCCS<-Results$Beta_Estimates
Final_Beta_LCCS$Method<-rep("Local case control sampling",nrow(Final_Beta_LCCS))
colnames(Final_Beta_LCCS)<-c("Method","Subsample",paste0("Beta",0:3))
```

## A- and L-optimality subsampling

```{r A- and L-optimality subsampling}
# A- and L-optimality subsampling for GLM 
NeEDS4BigData::ALoptimalGLMSub(r1=300,r2=rep_k,
                               Y=as.matrix(Original_Data[,1]),
                               X=as.matrix(Original_Data[,-1]),
                               N=N,family = "logistic")->Results
Final_Beta_ALoptGLM<-Results$Beta_Estimates
colnames(Final_Beta_ALoptGLM)<-c("Method","Subsample",paste0("Beta",0:3))
```

## A-optimality subsampling under measurement constraints

```{r A-optimality subsampling under measurement constraints}
# A-optimality subsampling for without response
NeEDS4BigData::AoptimalMCGLMSub(r1=300,r2=rep_k,
                                Y=as.matrix(Original_Data[,1]),
                                X=as.matrix(Original_Data[,-1]),
                                N=N,family="logistic")->Results
Final_Beta_AoptMCGLM<-Results$Beta_Estimates
Final_Beta_AoptMCGLM$Method<-rep("A-Optimality MC",nrow(Final_Beta_AoptMCGLM))
colnames(Final_Beta_AoptMCGLM)<-c("Method","Subsample",paste0("Beta",0:3))
```

## Summary

```{r summarising results}
# Summarising Results of Scenario one
Final_Beta<-rbind(Final_Beta_RS,Final_Beta_LS,Final_Beta_LCCS,
                  Final_Beta_ALoptGLM,Final_Beta_AoptMCGLM)

# Obtaining the model parameter estimates for the full big data model
glm(Y~.-1,data = Original_Data,family="binomial")->All_Results
coefficients(All_Results)->All_Beta
matrix(rep(All_Beta,by=nrow(Final_Beta)),nrow = nrow(Final_Beta),
       ncol = ncol(Final_Beta[,-c(1,2)]),byrow = TRUE)->All_Beta

remove(Final_Beta_RS,Final_Beta_LS,Final_Beta_LCCS,Final_Beta_ALoptGLM,
       Final_Beta_AoptMCGLM,Temp_Data,Results,All_Results)
```

The mean squared error is plotted below.

```{r Model parameter estimates,fig.width=7,fig.height=12,fig.align='center',fig.cap="Mean squared error for a) all the subsampling methods and b) without unweighted leverage sampling."}
# Obtain the mean squared error for the model parameter estimates
MSE_Beta<-data.frame("Method"=Final_Beta$Method,
                     "Subsample"=Final_Beta$Subsample,
                     "MSE"=rowSums((All_Beta - Final_Beta[,-c(1,2)])^2)) 

MSE_Beta$Method<-factor(MSE_Beta$Method,levels = Method_Names,labels = Method_Names)

# Plot for the mean squared error with all methods
ggplot(MSE_Beta |> dplyr::group_by(Method,Subsample) |> 
       dplyr::summarise(MSE=mean(MSE), .groups = 'drop'),
    aes(x=factor(Subsample),y=MSE,color=Method,group=Method,
        linetype=Method,shape=Method))+
  geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
  scale_color_manual(values = Method_Color)+
  scale_linetype_manual(values=Method_Line_Types)+
  scale_shape_manual(values = Method_Shape_Types)+
  theme_bw()+guides(colour = guide_legend(nrow = 3))+Theme_special()->p1

# Plot for the mean squared error with all methods except unweighted leverage
ggplot(MSE_Beta[!(MSE_Beta$Method %in% c("Unweighted Leverage","Shrinkage Leverage",
                                         "Basic Leverage")),] |> 
       dplyr::group_by(Method,Subsample) |> 
       dplyr::summarise(MSE=mean(MSE),.groups = 'drop'),
    aes(x=factor(Subsample),y=MSE,color=Method,group=Method,
        linetype=Method,shape=Method))+
  geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
  scale_color_manual(values = Method_Color[-7])+
  scale_linetype_manual(values=Method_Line_Types[-7])+
  scale_shape_manual(values = Method_Shape_Types[-7])+
  theme_bw()+guides(colour = guide_legend(nrow = 3))+Theme_special()->p2

ggarrange(p1,p2,nrow = 2,ncol = 1,labels = "auto")
```

### THANK YOU
