---
title: "Model based Subsampling for One Million Songs Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model based Subsampling for One Million Songs Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: Vignette_Ref.bib
---

<style>
body {
text-align: justify}
</style>

```{r knitr options,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages and theme,message=FALSE}
# Load the R packages
library(NeEDS4BigData)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(tidyr)
library(psych)
library(gam)

# Theme for plots
Theme_special<-function(){
  theme(legend.key.width= unit(1,"cm"),
        axis.text.x= element_text(color= "black", size= 12, angle= 30, hjust=0.75),
        axis.text.y= element_text(color= "black", size= 12), 
        strip.text= element_text(colour= "black", size= 12, face= "bold"),
        panel.grid.minor.x= element_blank(),
        axis.title= element_text(color= "black", face= "bold", size= 12),
        legend.text= element_text(color= "black", size= 11),
        legend.position= "bottom",
        legend.margin= margin(0,0,0,0),
        legend.box.margin= margin(-1,-2,-1,-2)) 
}
```

# Understanding the one million songs data

Echo nest collected data to understand songs and their popularity using their play counts over 1 million users based on the songs basic information, such as duration, tempo, loudness, etc.
This data was transformed accordingly as in the works @bertin2011million to study the number of counts through the songs features. 
This data contains $6$ columns and has $205,032$ observations, first column is the response variable and the rest are covariates. 
We consider the covariates a) duration ($x_1$), b) loudness ($x_2$), c) tempo ($x_3$), d) artist hotness ($x_4$), and e) song hotness ($x_5$) to model the response, the number of play counts for each song.
The covariates $x_1,x_2,x_3$ are scaled to be mean of zero and variance of one, while the remaining covariates are in between $0-1$.
For the given data subsampling methods are implemented assuming the main effects model can describe the data.

```{r Exploring data}
# Selecting 100% of the big data and prepare it
Original_Data<-One_Million_Songs
colnames(Original_Data)<-c("Y",paste0("X",1:ncol(Original_Data[,-1])))

# Scaling the covariate data
for (j in 2:4) {
  Original_Data[,j]<-scale(Original_Data[,j])
}

head(One_Million_Songs) %>% 
  kable(format = "html",
        caption = "First few observations of the one million songs data.")

# Setting the sample sizes
N<-nrow(Original_Data); M<-250; k<-seq(8,20,by=2)*100; rep_k<-rep(k,each=M)
```

Based on this model the methods random sampling, leverage sampling [@ma2014statistical;@ma2015leveraging], $A$- and $L$-optimality subsampling [@ai2021optimal;@yao2021review] and $A$-optimality subsampling with response constraint [@zhang2021optimal] are implemented on the big data.

```{r Define the methods}
# define colors, shapes, line types for the methods
Method_Names<-c("A-Optimality","L-Optimality","A-Optimality MC",
                "Shrinkage Leverage","Basic Leverage","Unweighted Leverage",
                "Random Sampling")
Method_Colour<-c("#BBFFBB","#50FF50","#00BB00",
                 "#F76D5E","#D82632","#A50021","#000000")
Method_Shape_Types<-c(rep(17,3),rep(4,3),16)
```

The obtained samples and their respective model parameter estimates are over $M=250$ simulations across different sample sizes $k=(800,\ldots,2000)$. 
We set the initial sample size of $r1=400$ for the methods that requires a random sample. 
From the final samples, model parameters of the assume model are estimated.
These estimated model parameters are compared with the estimated model parameters of the full big data through the mean squared error $MSE(\tilde{\beta}_k,\hat{\beta})=\frac{1}{MJ}\sum_{i=1}^M \sum_{j=1}^J(\tilde{\beta}_{k,j} - \hat{\beta}_j)^2$. 
Here, $\tilde{\beta}_k$ is the estimated model parameters from the sample of size $k$ and $\hat{\beta}$ is the estimated model parameters from the full big data, while $j$ is index of the model parameter.

## Random sampling

Below is the code for this implementation.

```{r Random sampling}
# Random sampling
Final_Beta_RS<-matrix(nrow = length(k)*M,ncol = ncol(Original_Data[,-1])+1)
for (i in 1:length(rep_k)) {
  Temp_Data<-Original_Data[sample(1:N,rep_k[i]),]
  glm(Y~.-1,data=Temp_Data,family="poisson")->Results
  Final_Beta_RS[i,]<-c(rep_k[i],coefficients(Results))
  if(i==length(rep_k)){print("All simulations completed for random sampling")}
}
Final_Beta_RS<-cbind.data.frame("Random Sampling",Final_Beta_RS)
colnames(Final_Beta_RS)<-c("Method","Sample",paste0("Beta",1:5))
```

## Leverage sampling

```{r Leverage sampling}
# Leverage sampling 
## we set the shrinkage value of S_alpha=0.9
NeEDS4BigData::LeverageSampling(r=rep_k,Y=as.matrix(Original_Data[,1]),
                                X=as.matrix(Original_Data[,-1]),N=N,
                                S_alpha = 0.9,family = "poisson")->Results
Final_Beta_LS<-Results$Beta_Estimates
colnames(Final_Beta_LS)<-c("Method","Sample",paste0("Beta",1:5))
```

## A- and L-optimality subsampling

```{r A- and L-optimality subsampling}
# A- and L-optimality subsampling for GLM
NeEDS4BigData::ALoptimalGLMSub(r1=400,r2=rep_k,
                               Y=as.matrix(Original_Data[,1]),
                               X=as.matrix(Original_Data[,-1]),
                               N=N,family = "poisson")->Results
Final_Beta_ALoptGLM<-Results$Beta_Estimates
colnames(Final_Beta_ALoptGLM)<-c("Method","Sample",paste0("Beta",1:5))
```

## A-optimality subsampling with measurement constraints

```{r A-optimality subsampling with measurement constraints}
# A-optimality subsampling without response
NeEDS4BigData::AoptimalMCGLMSub(r1=400,r2=rep_k,
                                Y=as.matrix(Original_Data[,1]),
                                X=as.matrix(Original_Data[,-1]),
                                N=N,family="poisson")->Results
Final_Beta_AoptMCGLM<-Results$Beta_Estimates
Final_Beta_AoptMCGLM$Method<-rep("A-Optimality MC",nrow(Final_Beta_AoptMCGLM))
colnames(Final_Beta_AoptMCGLM)<-c("Method","Sample",paste0("Beta",1:5))
```

## Summary

```{r summarising results}
# Summarising Results of Scenario one
Final_Beta<-rbind(Final_Beta_RS,Final_Beta_LS,
                  Final_Beta_ALoptGLM,Final_Beta_AoptMCGLM)

# Obtaining the model parameter estimates for the full big data model
glm(Y~.-1,data = Original_Data,family="poisson")->All_Results
coefficients(All_Results)->All_Beta
matrix(rep(All_Beta,by=nrow(Final_Beta)),nrow = nrow(Final_Beta),
       ncol = ncol(Final_Beta[,-c(1,2)]),byrow = TRUE)->All_Beta

remove(Final_Beta_RS,Final_Beta_LS,Final_Beta_ALoptGLM,
       Final_Beta_AoptMCGLM,Temp_Data,Results,All_Results)
```

The mean squared error is plotted below.

```{r First scenario model parameter estimates,fig.width=7,fig.height=10,fig.align='center',fig.cap="Mean squared error for a) all the subsampling methods and b) without unweighted leverage sampling, with 5% and 95% percentile intervals."}
# Obtain the mean squared error for the model parameter estimates
MSE_Beta<-data.frame("Method"=Final_Beta$Method,
                     "Sample"=Final_Beta$Sample,
                     "SE"=rowSums((All_Beta - Final_Beta[,-c(1,2)])^2)) 

MSE_Beta$Method<-factor(MSE_Beta$Method,levels = Method_Names,labels = Method_Names)

Mean_Data <- MSE_Beta |>
    group_by(Method,Sample) |>
    dplyr::summarise(Mean = mean(SE),
                     min=quantile(SE,0.05),
                     max=quantile(SE,0.95), .groups = "drop")

# Plot for the mean squared error with all methods
ggplot(data=Mean_Data,aes(x=factor(Sample),y=Mean,color=Method,shape=Method))+
  xlab("Sample Size")+ylab("Squared Error")+
  geom_point(size=3,position = position_dodge(width = 0.5))+
  geom_line(data=Mean_Data,aes(x=factor(Sample),y=Mean,group=Method,color=Method),
            position = position_dodge(width = 0.5))+
  geom_errorbar(data=Mean_Data,aes(ymin=min,ymax=max),
                width=0.3,position = position_dodge(width = 0.5))+
  scale_color_manual(values = Method_Colour)+
  scale_shape_manual(values = Method_Shape_Types)+
  theme_bw()+Theme_special()+
  guides(colour = guide_legend(nrow = 3))->p1

# Plot for the mean squared error with all methods except unweighted leverage
ggplot(data=Mean_Data[Mean_Data$Method != "Unweighted Leverage",],
       aes(x=factor(Sample),y=Mean,color=Method,shape=Method))+
  xlab("Sample Size")+ylab("Squared Error")+
  geom_point(size=3,position = position_dodge(width = 0.5))+
  geom_line(data=Mean_Data[Mean_Data$Method != "Unweighted Leverage",],
            aes(x=factor(Sample),y=Mean,group=Method,color=Method),
            position = position_dodge(width = 0.5))+
  geom_errorbar(data=Mean_Data[Mean_Data$Method != "Unweighted Leverage",],
                aes(ymin=min,ymax=max),width=0.3,
                position = position_dodge(width = 0.5))+
  scale_color_manual(values = Method_Colour[-7])+
  scale_shape_manual(values = Method_Shape_Types[-7])+
  theme_bw()+Theme_special()+
  guides(colour = guide_legend(nrow = 3))->p2

ggarrange(p1,p2,nrow = 2,ncol = 1,labels = "auto")
```

## References
