---
title: "Poisson Regression : Model robust or misspecification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Poisson Regression : Model robust or misspecification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: Vig_Method_Ref.bib  
---

<style>
body {
text-align: justify}
</style>

```{r load packages and theme,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NeEDS4BigData)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(tidyr)
library(psych)
library(gam)

Theme_special<-function(){
  theme(legend.key.width=unit(1,"cm"),
        axis.text.x = element_text(color = "black",size=12, angle = 30, hjust=0.75), 
        axis.text.y = element_text(color = "black",size=12),
        strip.text = element_text(colour = "black",size = 12,face="bold"),
        panel.grid.minor.x = element_blank(),
        axis.title= element_text(color = "black",face = "bold",size = 12),
        legend.text = element_text(color = "black", size=11),
        legend.position = "bottom",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-1,-2,-1,-2)) 
}
```

# Understanding the bike sharing data

The data contains $4$ columns and has $17,379$ observations, first column is the response variable and the rest are covariates. 
We consider the covariates a) temperature ($x_1$), b) humidity ($x_2$) and c) windspeed ($x_3$) to model the response, the number of bikes rented hourly.
The covariates are scaled to be mean of $\mu=0$ and variance of $\sigma^2=1$.

Given data is analysed under two different scenarios,

1. model robust or averaging subsampling methods assuming that a set of models can describe the data.
2. subsampling method assuming the main effects model is potentially misspecified.

```{r Exploring data}
Original_Data<-cbind(Bike_sharing[,1],1,Bike_sharing[,-1])
colnames(Original_Data)<-c("Y",paste0("X",0:ncol(Original_Data[,-c(1,2)])))

# Scaling the covariate data
for (j in 3:5) {
  Original_Data[,j]<-scale(Original_Data[,j])
}
head(Original_Data) %>% 
  kable(format = "html",
        caption = "First five observations of the bike sharing data.")

# Setting the subsample sizes
N<-nrow(Original_Data); M<-200; k<-c(6:15)*100; rep_k<-rep(k,each=M)
```

## Model robust or model averaging subsampling

The method $A$- and $L$-optimality of model robust or average subsampling is compared against the $A$- and $L$-optimality subsampling method. 
Here five different models are considered 1) main effects model ($\beta_0+\beta_1X_1+\beta_2X_2+\beta_3X_3$), 2-4) main effects model with the squared term of each covariate ($X^2_1 / X^2_2 / X^2_3$) and 5) main effects model with all the squared terms ($\beta_0+\beta_1X_1+\beta_2X_2+\beta_3X_3+\beta_4X^2_1+\beta_5X^2_2+\beta_6X^2_3$). 
For each model $j$ the mean squared error of the model parameters $MSE_j(\tilde{\beta}_k,\hat{\beta})=\sum_{i=1}^M (\tilde{\beta}_k - \hat{\beta})^2/M$ are calculated for the $M=100$ simulations across the subsample sizes $k=(600,\ldots,1500)$ and the initial subsample size is $r1=300$. 
Here, for the $j$-th model $\tilde{\beta}_k$ is the estimated model parameters from the subsample of size $k$ and $\hat{\beta}$ is the estimated model parameters from the full big data.

```{r define subsampling methods}
# Define the subsampling methods and their respective colors, shapes and line types
Method_Names<-c("A-Optimality","L-Optimality","A-Optimality MR","L-Optimality MR")
Method_Color<-c("red","darkred","palegreen","springgreen4")
Method_Shape_Types<-c(rep(8,2),rep(4,2))
Method_Line_Types<-c(rep("twodash",2),rep("dotted",2))

# Preparing the data for the model average method with squared terms
No_of_Variables<-ncol(Original_Data[,-c(1,2)])
Squared_Terms<-paste0("X",1:No_of_Variables,"^2")
term_no <- 2
All_Models <- list(c("X0",paste0("X",1:No_of_Variables)))

Original_Data_ModelRobust<-cbind(Original_Data,Original_Data[,-c(1,2)]^2)
colnames(Original_Data_ModelRobust)<-c("Y","X0",paste0("X",1:No_of_Variables),
                                       paste0("X",1:No_of_Variables,"^2"))

for (i in 1:No_of_Variables)
{
  x <- as.vector(combn(Squared_Terms,i,simplify = FALSE))
  for(j in 1:length(x))
  {
    All_Models[[term_no]] <- c("X0",paste0("X",1:No_of_Variables),x[[j]])
    term_no <- term_no+1
  }
}
All_Models<-All_Models[-c(5:7)]
names(All_Models)<-paste0("Model_",1:length(All_Models))
```

### Apriori probabilities are equal 

Consider for $Q=5$ each model has an equal a priori probability (i.e $\alpha_q=1/5,q=1,\ldots,5$). 
Below is the code of implementation for this scenario.

```{r Equal apriori,fig.width=12,fig.height=12,fig.align='center',fig.cap="Mean squared error for all the models with equal apriori in the order a to e for Model 1 to 5 across the subsampling methods under comparison."}
# A- and L-optimality model robust subsampling for poisson regression
NeEDS4BigData::modelRobustPoiSub(r1=300,r2=rep_k,
                                 Y=as.matrix(Original_Data_ModelRobust[,1]),
                                 X=as.matrix(Original_Data_ModelRobust[,-1]),
                                 N=N,Alpha=rep(1/length(All_Models),length(All_Models)),
                                 All_Combinations = All_Models,
                                 All_Covariates = colnames(Original_Data_ModelRobust)[-1])->Results
Final_Beta_modelRobust<-Results$Beta_Estimates

# Mean squared error and their respective plots for all five models
MSE_Beta_MR<-list(); plot_list_MR<-list()
for (i in 1:length(All_Models)) {
  glm(Y~.-1,data=Original_Data_ModelRobust[,c("Y",All_Models[[i]])],family="poisson")->All_Results
  All_Beta<-coefficients(All_Results)

  matrix(rep(All_Beta,by=nrow(Final_Beta_modelRobust[[i]])),
         nrow = nrow(Final_Beta_modelRobust[[i]]),
         ncol = ncol(Final_Beta_modelRobust[[i]][,-c(1,2)]),byrow = TRUE)->All_Beta

  MSE_Beta_MR[[i]]<-data.frame("Method"=Final_Beta_modelRobust[[i]]$Method,
                               "Subsample"=Final_Beta_modelRobust[[i]]$r2,
                               "MSE"=rowSums((All_Beta - Final_Beta_modelRobust[[i]][,-c(1,2)])^2))

  ggplot(MSE_Beta_MR[[i]] |> dplyr::group_by(Method,Subsample) |> 
           dplyr::summarise(MSE=mean(MSE), .groups = 'drop'),
         aes(x=factor(Subsample),y=MSE,color=Method,group=Method,linetype=Method,shape=Method))+
    geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
    scale_color_manual(values = Method_Color)+
    scale_linetype_manual(values=Method_Line_Types)+
    scale_shape_manual(values = Method_Shape_Types)+
    theme_bw()+guides(colour = guide_legend(nrow = 2))+Theme_special()->plot_list_MR[[i]]
}

ggarrange(plotlist = plot_list_MR,nrow = 3,ncol = 2,labels = "auto",
          legend = "bottom",common.legend = TRUE)
```

### Apriori probabilities are unequal 

Consider for $Q=5$ each model has a different a priori probability that is based on the number of model parameters as in [@scott2010bayes] (i.e., $\alpha_q = \Big(7 {6 \choose b_q}\Big)^{−1}$ where $b_q$ is the number of model parameters in the $q$-th model excluding the intercept).
Below is the code of implementation for this scenario.

```{r Unequal apriori, fig.width=12,fig.height=12,fig.align='center',fig.cap="Mean squared error for all the models with unequal apriori in the order a to e for Model 1 to 5 across the subsampling methods under comparison."}
UnEqual_Apriori<-NULL
for (i in 1:length(All_Models)) 
{
  UnEqual_Apriori[i]<-(7*choose(6,length(All_Models[[i]])-1))^(-1)
}

UnEqual_Apriori<-UnEqual_Apriori/sum(UnEqual_Apriori)

# A- and L-optimality model robust subsampling for poisson regression
NeEDS4BigData::modelRobustPoiSub(r1=300,r2=rep_k,
                                 Y=as.matrix(Original_Data_ModelRobust[,1]),
                                 X=as.matrix(Original_Data_ModelRobust[,-1]),
                                 N=N,Alpha=UnEqual_Apriori,
                                 All_Combinations = All_Models,
                                 All_Covariates = colnames(Original_Data_ModelRobust)[-1])->Results
Final_Beta_modelRobust<-Results$Beta_Estimates

# Mean squared error and their respective plots for all five models
MSE_Beta_MR<-list(); plot_list_MR<-list()
for (i in 1:length(All_Models)) {
  glm(Y~.-1,data=Original_Data_ModelRobust[,c("Y",All_Models[[i]])],
      family = "poisson")->All_Results
  All_Beta<-coefficients(All_Results)
  
  matrix(rep(All_Beta,by=nrow(Final_Beta_modelRobust[[i]])),
         nrow = nrow(Final_Beta_modelRobust[[i]]),
         ncol = ncol(Final_Beta_modelRobust[[i]][,-c(1,2)]),byrow = TRUE)->All_Beta
  
  MSE_Beta_MR[[i]]<-data.frame("Method"=Final_Beta_modelRobust[[i]]$Method,
                               "Subsample"=Final_Beta_modelRobust[[i]]$r2,
                               "MSE"=rowSums((All_Beta - Final_Beta_modelRobust[[i]][,-c(1,2)])^2))
  
  ggplot(MSE_Beta_MR[[i]] |> dplyr::group_by(Method,Subsample) |> 
           dplyr::summarise(MSE=mean(MSE),.groups ='drop'),
         aes(x=factor(Subsample),y=MSE,color=Method,group=Method,linetype=Method,shape=Method))+
    geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
    scale_color_manual(values = Method_Color)+
    ggtitle(paste0("Model ",i))+
    scale_linetype_manual(values=Method_Line_Types)+
    scale_shape_manual(values = Method_Shape_Types)+
    theme_bw()+guides(colour = guide_legend(nrow = 2))+Theme_special()->plot_list_MR[[i]]
}

ggarrange(plotlist = plot_list_MR,nrow = 3,ncol = 2,labels = "auto",
          common.legend = TRUE,legend = "bottom")
```

## Main effects model is potentially misspecified

The final and third scenario is for comparison of the subsampling method under the assumption that the main effects model is potentially misspecified against the $A$- and $L$-optimality subsampling method. 
Under the subsampling method that accounts for potential model misspecification we take the scaling factor of $\alpha=10$. 
As in scenario one an two the number of simulations and the subsample sizes stay the same. 
We compare the mean squared error of the estimated model parameters, however as we assume the model is potentially misspecified the asymptotic approximation of the mean squared error from the predictions are calculated as well.
Below is the code for this implementation.

```{r Potential model misspecification methods,fig.width=6,fig.height=6,fig.align='center', warning=FALSE,fig.cap="Average loss for the potentially misspecified main effects model across the subsampling methods under comparison."}
# Define the subsampling methods and their respective colors, shapes and line types
Method_Names<-c("A-Optimality","L-Optimality","LmAMSE","LmAMSE Log Odds 10","LmAMSE Power 10")
Method_Color<-c("red","darkred","palegreen","green","springgreen4")
Method_Shape_Types<-c(rep(8,2),rep(4,3))
Method_Line_Types<-c(rep("twodash",2),rep("dotted",3))

# For the big data fit the main effects model and estimate the contamination
interaction_terms <- combn(colnames(Original_Data[,-1])[-1],2,FUN=function(x)paste(x,collapse="*"))
my_formula<-as.formula(paste("Y~",paste(paste0("s(X",1:ncol(Original_Data[,-c(1,2)]),")"),collapse="+"),
                             "+",paste(paste0("s(",interaction_terms,")"),collapse=" + ")))
glm(Y~.-1,data=Original_Data,family="poisson")->Results
beta.prop<-coefficients(Results)
Xbeta_Final<-as.vector(as.matrix(Original_Data[,-1])%*%beta.prop)
fit_GAM<-gam::gam(formula = my_formula,data=Original_Data,family="poisson")
Xbeta_GAM<-gam::predict.Gam(fit_GAM,newdata = Original_Data[,-1])
f_estimate<-Xbeta_GAM - Xbeta_Final

# A- and L-optimality and LmAMSE model misspecified subsampling for poisson regression 
NeEDS4BigData::modelMissPoiSub(r1=300,r2=rep_k,
                               Y=as.matrix(Original_Data[,1]),
                               X=as.matrix(Original_Data[,-1]),
                               N=N,Alpha=10, Beta_Estimate_Full = beta.prop,
                               F_Estimate_Full = f_estimate)->Results

Final_Beta_modelMiss<-Results$Beta_Estimates
Final_Loss_modelMiss<-Results$Loss_Estimates

matrix(rep(beta.prop,by=nrow(Final_Beta_modelMiss)),nrow = nrow(Final_Beta_modelMiss),
       ncol = ncol(Final_Beta_modelMiss[,-c(1,2)]),byrow = TRUE)->All_Beta

# Plots for the mean squared error of the model parameter estimates 
# and the loss for the main effects model 
MSE_Beta_modelMiss<-data.frame("Method"=Final_Beta_modelMiss$Method,
                               "Subsample"=Final_Beta_modelMiss$r2,
                               "MSE"=rowSums((All_Beta - Final_Beta_modelMiss[,-c(1,2)])^2)) 

ggplot(MSE_Beta_modelMiss |> dplyr::group_by(Method,Subsample) |> 
         dplyr::summarise(MSE=mean(MSE), .groups = 'drop'),
       aes(x=factor(Subsample),y=MSE,color=Method,group=Method,linetype=Method,shape=Method))+
  geom_point()+geom_line()+xlab("Subsample size")+ylab("MSE")+
  scale_color_manual(values = Method_Color)+
  scale_linetype_manual(values=Method_Line_Types)+
  scale_shape_manual(values = Method_Shape_Types)+
  theme_bw()+guides(colour = guide_legend(nrow = 3))+Theme_special()->p1

ggplot(Final_Loss_modelMiss |> dplyr::group_by(r2,Method) |> 
         dplyr::summarise(meanLoss=mean(Loss), .groups = 'drop'),
       aes(x=factor(r2),y=meanLoss,color=Method,group=Method,linetype=Method,shape=Method)) +
  geom_point()+geom_line()+xlab("Subsample size")+ylab("Mean Loss")+
  scale_color_manual(values = Method_Color)+
  scale_linetype_manual(values=Method_Line_Types)+
  scale_shape_manual(values = Method_Shape_Types)+
  theme_bw()+guides(colour = guide_legend(nrow = 3))+Theme_special()->p2

# ggarrange(p1,p2,nrow = 2,ncol = 1,labels = "auto",
#           common.legend = TRUE,legend = "bottom")
p2
```

### THANK YOU 
